Supplement to Global Distribution of Unintended Pregnancies and Contraceptive Usage Project

### 2024-04-29 code/analysis-code/statistical-analysis.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

---
title: "Statistical Analysis"
author: "Taylor Glass and Arlyn Santiago"
date: "03/12/24"
output: html_document
---
## Load necessary packages
```{r}
#| message: false
library(dplyr) # for basic syntax
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) # for model building 
library(yardstick) # for finding performance metrics 
library(rsample) #for splitting the data into testing/training
library(knitr) # to create performance metric tables 
library(parsnip) # to create a workflow for cross-validation
library(glmnet) # for LASSO regression
library(ranger) # for random forest model 
```

@@

# create a professional format of the table to add to the manuscript 
kable_lmtable1 <- kable(lmtable1)

# save fit results table  
table_file1 = here("results", "tables", "resulttable1.rds")
saveRDS(kable_lmtable1, file = table_file1)
```
There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.

@@

# create a professional format of the table to add to the manuscript 
kable_lmtable2 <- kable(lmtable2)

# save fit results table  
table_file2 = here("results", "tables", "resulttable2.rds")
saveRDS(kable_lmtable2, file = table_file2)

@@

# create a professional format of the table to add to the manuscript 
kable_lmtable8 <- kable(lmtable8)

# save fit results table  
table_file8 = here("results", "tables", "resulttable8.rds")
saveRDS(kable_lmtable8, file = table_file8)
```
This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.

>>>>

---
title: "Statistical Analysis"
author: "Taylor Glass and Arlyn Santiago"
date: "03/12/24"
output: html_document
---
## Load necessary packages
```{r}
#| message: false
library(dplyr) # for basic syntax
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) # for model building 
library(flextable) # for creating tables for the manuscript
library(yardstick) # for finding performance metrics 
library(rsample) #for splitting the data into testing/training
library(knitr) # to create performance metric tables 
library(parsnip) # to create a workflow for cross-validation
library(glmnet) # for LASSO regression
library(ranger) # for random forest model 
```

@@

# create a professional format of the table to add to the manuscript 
lmtable1 <- as_flextable(lmtable1)

# save fit results table  
table_file1 = here("results", "tables", "resulttable1.rds")
saveRDS(lmtable1, file = table_file1)
```
There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.

@@

# create a professional format of the table to add to the manuscript 
lmtable2 <- as_flextable(lmtable2)

# save fit results table  
table_file2 = here("results", "tables", "resulttable2.rds")
saveRDS(lmtable2, file = table_file2)

@@

# create a professional format of the table to add to the manuscript 
lmtable8 <- as_flextable(lmtable8)

# save fit results table  
table_file8 = here("results", "tables", "resulttable8.rds")
saveRDS(lmtable8, file = table_file8)
```
This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.

### 2024-04-29 code/eda-code/eda.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
plot_3 <- mydata %>% ggplot(aes(x=sarc_standard, y=pct_upreg_sarc)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=plot_3)

@@

# explore percent unintended pregnancies using short acting reversible methods by region
plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_jitter(aes(colour = region)) +
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=plot_4)

>>>>

Plot 3 uses `pct_upreg_sarc` and `sarc_standard` to examine the relationship with short-acting reversible methods. 
```{r}
# explore percentage of unintended pregnancies compared to standardized count of women using short acting reversible methods
plot_3 <- mydata %>% ggplot(aes(x=sarc_standard, y=pct_upreg_sarc)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardized Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_3)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png")
ggsave(filename = figure_file, plot=plot_3)

@@

# explore percent unintended pregnancies using short acting reversible methods by region
plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_jitter(aes(colour = region)) +
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardized Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=plot_4)

### 2024-04-29 products/manuscript/Manuscript.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

```{r, echo=FALSE, message=FALSE}
# load a few R packages for rendering 
library(here)
library(readr)
library(knitr)
library(dplyr)
library(naniar)
library(ggplot2)
```

@@

Step three is the statistical analysis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder. We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion. We also explore the differences in the outcome between regions using an ANOVA test and Tukey's Honestly Significant Difference test for pairwise comparisons within ANOVA. Result tables for each model and test are stored in the results > tables folder.

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. This process is similar to cross-validation because it relies on resampling the training data. Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We use  a grid search to ensure that we try all combinations of the parameters in our grid. We complete the tuning process with a 5-fold cross-validation repeated 5 times. We added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so. Each of these processes are crucial in determining which model performs best for our dataset. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1. [insert explanation here]
```{r}
#| message: false
#| label: tbl-1
#| tbl-cap: "[A summary table was done to present data about the seven predictor variables chosen in a structured format. Summary tables enable us to compare between different categories or groups within the data set. This can reveal trends or relationships. From the summary table below, it would appear that the region with the highest rate of unintended pregnancies due to SARC usage is Latin America/Caribbean at 24%, followed by Europe at 23%. Africa and Asia have a mean of 13% and 14% respectively, while Oceania appears to have the lowest rate at 6%. It appears that a higher percentage of women are currently married in Asia (66%) and Africa (60%) compared to Latin America/Caribbean (54%) and Oceania (60%).The average percentage of never married women across the 5 regions is 31%. Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.]"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "[A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods. The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "[A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies. In this case, a positive correlation is shown. It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "[Scatter plots are used to see if there is a relationship between two sets of continuous data. In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies. The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "[Violin plots are used to visualize the distribution of data across different categories. In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage. This means that there is more variation in unintended pregnancy rates in counties with high SARC usage.]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

@@

```{r}
#| message: false
#| label: tbl-2
#| fig-cap: "This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception. This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome. The coefficient of 87.65998 is the largest single coefficient of any predictor that we studied, which increases our confidence in including it as a predictor in our final model."
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @tbl-3.
```{r}
#| message: false
#| label: tbl-3
#| fig-cap: "In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia. Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome. This model chose the African region as the reference group, which means that it has the weakest association with the outcome compared to the other 4 regions."
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 10 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

>>>>

```{r, echo=FALSE, message=FALSE}
# load a few R packages for rendering 
library(here) # for locating tables and figures
library(readr) # for loading tables and figures
library(knitr) # for displaying tables and figures 
library(dplyr) # for basic syntax
```

@@

Step three is the statistical analysis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder. We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion. Result tables for each model and test are stored in the results > tables folder.

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). Once we decided on the final model based on RMSE and MAE, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. Next, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We used a grid search to ensure that we try all combinations of the parameters in our grid. Lastly, we added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. Each of these processes are crucial in determining which model performs best for our data. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, bootstrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1. A summary table was done to present data about the seven predictor variables chosen in a structured format. Summary tables enable us to compare between different categories or groups within the data set. This can reveal trends or relationships. From the summary table below, it would appear that the region with the highest rate of unintended pregnancies due to SARC usage is Latin America/Caribbean at 24%, followed by Europe at 23%. Africa and Asia have a mean of 13% and 14% respectively, while Oceania appears to have the lowest rate at 6%. It appears that a higher percentage of women are currently married in Asia (66%) and Africa (60%) compared to Latin America/Caribbean (54%) and Oceania (60%).The average percentage of never married women across the 5 regions is 31%. Latin America/Caribbean region has the highest percentage of never married women at 35% and Asia region has has the lowest percentage of never married women at 29%.
```{r}
#| message: false
#| label: tbl-1
#| tbl-cap: "Table 1: Summary of All Explored Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "A bar chart was made to showcase the frequencies of the number of countries and the percentage of unintended pregnancy rates using SARC methods. The data presented suggests that there are fewer countries with a higher number of unintended pregnancies using SARCs considering more than half of the countries experience less than 20% of unintended pregnancies when using SARCs."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "A Scatterplot was contructed to see if there is a correlation between a standardized count of women using SARCs and the percentage of unintended pregnancies. In this case, a positive correlation is shown. It suggests that as the number of women using SARCs increases (x-axis), the percentage of unintended pregnancies increases (y-axis)"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "Scatter plots are used to see if there is a relationship between two sets of continuous data. In this scatter plot, the relationship between the standardized count of women using SARCs and the percentage of unintended pregnancies is being examined.There appears to be a positive correlation between the standardized count of women using SARCs and the percentage of unintended pregnancies. The regions with the highest percentage of unintended pregnacies are shown to be Latin America & the Caribbean, followed by Africa."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "Violin plots are used to visualize the distribution of data across different categories. In this case, it is being used to see the distribution of unintended pregnancy rates among countries that use short-acting reversible contraceptives (SARCs).The violin plot below shows that the distribution of unintended pregnancy rates is wider for countries with high SARC usage compared to counties with medium or low SARC usage. This means that there is more variation in unintended pregnancy rates in counties with high SARC usage."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

@@

```{r}
#| message: false
#| label: tbl-2
#| fig-cap: "Table 2: This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception. This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome. The coefficient of 87.65998 is the largest single coefficient of any predictor that we studied, which increases our confidence in including it as a predictor in our final model."
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @tbl-3.
```{r}
#| message: false
#| label: tbl-3
#| fig-cap: " Table 3: In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia. Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome. This model chose the African region as the reference group, which means that it has the weakest association with the outcome compared to the other 4 regions."
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 10 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

### 2024-04-29 products/manuscript/supplement/Supplementary-Material.qmd
Part 6 Updates
Used flextable to fix the table headings in the rendered document. Explained how to reproduce in the master READMe file. Responded to Cassia's feedback by fixing typos and deleting extra files.

We only briefly mentioned that we used cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript. Cross-validation was chosen as a method for determining how well the model performs on data it has not seen before because that is the main point of building models. Our original models, found in resulttables1-10 in the results > tables folder, are created using all of the data. The cross-validation process built models with 75% of the data that was designated as the training data, and the remaining 25% was used as testing data. Our dataset is small with only 132 observations, so the insights provided here are not robust. However, we still want to know how well these variables predict our outcome to determine how meaningful each predictor is, so we completed cross-validation anyways to answer this question. We also used this approach due to the risk of overfitting in our model based on interaction between variables. Plotting the observed and predicted values for our model allowed us to look for systematic deviations from the expected relationship visualized with a 45 degree line as shown in @supplement-1. Since we did conclude that the model overfits the data based on unhealthy deviations of the testing data from the 45 degree line, our use of cross-validation is appropriate and useful. Lastly, this approach is an easily interpretable way to increase confidence in the performance of our models based on metrics because both proceses lead to the same conclusions. 
```{r}
#| label: supplement-1
#| fig-cap: "Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```

@@

LASSO regression is important in this study because we need to manage the bias-variance tradeoff since we have evidence that our model overfits the data. By applying a penalty parameter to the cost function, LASSO regression shrinks the coefficients of the parameter estimates, so these estimates will always be smaller than the ones provided by the original linear model. As the penalty parameter increases, a larger penalty term is added to the SSR, which increases the value of the cost function. As the penalty increases, the RMSE value increases, and the performance of the model is considered worse and worse comparatively. We chose a parameter of 0.1 first because that is standard for LASSO regression, but we also completed a parameter tuning process to validate that choice. As shown in @supplement-3, the RMSE performance metric increases exponentially when the penalty parameter increases past 1, which validates our previous selection of 0.1 for the penalty parameter because that will produce the lowest RMSE possible. Since coefficients can be reduced all the way to zero, this analysis completes feature selection as well. Shrinking the coefficients also helps combat colinearity in variables, which we established between the percent of currently married women and percent of never married women, and the three rates of maternal mortality, respectively. LASSO models are also easier to interpet, which increases the validity of our findings.
```{r}
#| label: supplement-3
#| fig-cap: "Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

@@

# Additional data cleaning explanation
The first priority when cleaning the data was analyzing missing values, and our findings can be seen @tbl-summarytable. We found that the `pct_upreg` variable containing the percentage of unintended pregnancies is unreliable due to missing data that could not be explained. We decided to use the variables with percentage of unintended pregnancies related to each birth control method as our outcome variables because they are complete and were found directly based on the AIU protocol information. We printed the summary table here because it was useful in determining missingness and creating a new factor variable for the usage of each birth control method at the factor level later in the processing sequence. 
```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Summary table to discover missingness"
#| echo: FALSE
resulttable2 <- readRDS("../../results/tables/summarytable.rds")
kable(resulttable2)
```

>>>>

We chose two performance metrics, root mean square error (RMSE) and mean absolute error (MAE) because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. We only briefly mentioned that the next step is to use cross-validation, bootstrapping, and LASSO regression to test our models in the methods section of our manuscript. Cross-validation was chosen as a method for determining how well the model performs on data it has not seen before because that is the main point of building models. Our original models, found in resulttables1-10 in the results > tables folder, are created using all of the data. The cross-validation process built models with 75% of the data that was designated as the training data, and the remaining 25% was used as testing data. Our dataset is small with only 132 observations, so the insights provided here are not robust. However, we still want to know how well these variables predict our outcome to determine how meaningful each predictor is, so we completed cross-validation anyways to answer this question. We also used this approach due to the risk of overfitting in our model based on interaction between variables. Plotting the observed and predicted values for our model allowed us to look for systematic deviations from the expected relationship visualized with a 45 degree line as shown in @supplement-1. Since we did conclude that the model overfits the data based on unhealthy deviations of the testing data from the 45 degree line, our use of cross-validation is appropriate and useful. Lastly, this approach is an easily interpretable way to increase confidence in the performance of our models based on metrics because both proceses lead to the same conclusions. 
```{r}
#| label: supplement-1
#| fig-cap: "Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```

@@

LASSO regression is important in this study because we need to manage the bias-variance tradeoff since we have evidence that our model overfits the data. By applying a penalty parameter to the cost function, LASSO regression shrinks the coefficients of the parameter estimates, so these estimates will always be smaller than the ones provided by the original linear model. As the penalty parameter increases, a larger penalty term is added to the SSR, which increases the value of the cost function. As the penalty increases, the RMSE value increases, and the performance of the model is considered worse and worse comparatively. We chose a parameter of 0.1 first because that is standard for LASSO regression, but we also completed a parameter tuning process to validate that choice.  We complete the tuning process with a 5-fold cross-validation repeated 5 times. As shown in @supplement-3, the RMSE performance metric increases exponentially when the penalty parameter increases past 1, which validates our previous selection of 0.1 for the penalty parameter because that will produce the lowest RMSE possible. Since coefficients can be reduced all the way to zero, this analysis completes feature selection as well. Shrinking the coefficients also helps combat colinearity in variables, which we established between the percent of currently married women and percent of never married women, and the three rates of maternal mortality, respectively. LASSO models are also easier to interpet, which increases the validity of our findings.
```{r}
#| label: supplement-3
#| fig-cap: "Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

@@
